Pleiotropy of expression quantitative trait loci increases with the number of regulated genes, active cell types and tissues and bound transcription factors

### 2022-11-19 ms/ms02_intro.tex
NEH
MS

% GWAS in general
Genome-wide association studies (GWAS) have allowed the association of frequent genetic variants with common traits.
%
The number of GWAS has rapidly increased since the first GWAS around 15 years ago and has resulted in comprehensive databases such as the NHGRI-EBI GWAS Catalog \citep{2007...Nature,2018.Parkinson.Buniello}.
%
For instance, in 2019, NHGRI-EBI GWAS Catalog contained 5687 GWAS comprising 71,673 variant-trait associations from 3,567 publications \citep{2018.Parkinson.Buniello}.

@@

In this work, I have taken advantage of two large eQTL and GWAS data sets from the EBI eQTL Catalogue and the IEU OpenGWAS, respectively,
to develop a global colocalization analysis based on 127 eQTL studies and 418 GWAS across many cell types, tissues, and phenotypes.
%
I have used these eQTL/GWAS variants to investigate mechanisms of pleiotropic gene regulatory variants.
%
These eQTL/GWAS variants might be a useful resource for the annotation and prioritization of frequent variants.
%
Therefore, I have developed a web application to query the variants.

>>>>

% GWAS in general
Genome-wide association studies (GWAS) have allowed the association of frequent genetic variants with common traits.
%
The number of GWAS has rapidly increased since the first GWAS around 15 years ago and has resulted in comprehensive databases such as the NHGRI-EBI GWAS Catalog \citep{2007burton.barret,2018.Parkinson.Buniello}.
%
For instance, in 2019, NHGRI-EBI GWAS Catalog contained 5687 GWAS comprising 71,673 variant-trait associations from 3,567 publications \citep{2018.Parkinson.Buniello}.

@@

In this work, I have taken advantage of two large eQTL and GWAS data sets from the EBI eQTL Catalogue and the IEU OpenGWAS, respectively,
to develop a global colocalization analysis based on 127 eQTL studies and 418 GWAS across many cell types, tissues, and phenotypes.
%
I have used these eQTL/GWAS variants to investigate mechanisms of pleiotropic gene regulatory variants.
%
These eQTL/GWAS variants might be a useful resource for the annotation and prioritization of frequent variants.
%
Therefore, I have developed an easy-to-use web application to expose these colocalization data.

### 2022-11-19 ms/ms03_results.tex
MS

\subsection*{Which genomic regions are pleiotropic regarding eQTLs?}

Pleiotropic eQTLs are aggregated in genomic regions, for instance, in cytobands 3q23, 5q31.1, 9p21.3, 15q24.1, and 19q13.33 for the eQTLs involved on five or more classes (Supplementary table 3).
%
Therefore I computed regions that include a large number of pleiotropic eQTLs (See Methods) (Table \ref{tab:pleiotropic_regions} and Supplementary table 4).
%
I found 13 regions with 5 or more trait classes, 30 regions with 4 or more trait classes and 80 regions with 3 or more trait classes (Supplementary table 4).
%
75\% of regions involved in two or more trait classes are shorter than 100 kb, 15\% of regions are between 100 kb and 200 kb, and less than 5\% are larger than 200 kb (Fig. \ref{fig:pleiotropy_region_distribution}a).

@@

Very pleiotropic regions with 5 trait classes or more make less than 1 Mb of the genome, while moderate pleiotropic regions with 3 trait classes or more make round 3 Mb of the genome (Fig. \ref{fig:pleiotropy_region_distribution}b.

\subsection*{How specific to traits, genes and cell type/tissues? are colocalized eQTLs/GWAS variants}

>>>>

\subsection*{Which genomic regions contain the most pleiotropic eQTLs?}

Pleiotropic eQTLs are concentrated in genomic regions such as in cytobands 3q23,
5q31.1, 9p21.3, 15q24.1, and 19q13.33 for eQTLs involved on five or more classes (Supplementary table 3).
%
Therefore I computed regions that include at least one pleiotropic eQTL in
sliding windows of 100,000 nt (See Methods) (Table \ref{tab:pleiotropic_regions} and Supplementary table 4).
%
I found 13 regions with 5 or more trait classes, 30 regions with 4 or more trait
classes and 80 regions with 3 or more trait classes (Supplementary table 4).
%
50\% of regions with at least two trait classes are shorter than 100 kb, 20\% of regions are
between 100 kb and 200 kb, 10\% between 200 kb and 300 kb, and 20\% are larger than 300 kb (Fig. \ref{fig:pleiotropy_region_distribution}a).

@@

Pleiotropic regions with 3 or more trait classes make around 3 Mb of the genome (Fig. \ref{fig:pleiotropy_region_distribution}b).
Very pleiotropic regions with 5 trait classes or more make around 0.6 Mb of the genome,
regions with 4 GWAS classes make an additional 1 Mb and regions with 3 GWAS classes make an additional 1.5 Mb (Fig. \ref{fig:pleiotropy_region_distribution}b).

\subsection*{How specific are pleiotropic eQTLs to traits, genes and tissues?}
